---
title: "Relationship between MPG and Other Variables for mtcars"
author: "Klevin D."
date: "April 27, 2016"
output:
    html_document:
        fig_caption: true
---
##Executive Summary
Motor Trend, a magazine about the automobile industry is interested in looking at a data set of a collection of cars. In this project, we will explore the relationship between a set of variables and miles per gallon (MPG) (outcome). The following two questions are for particular interest:

1. “Is an automatic or manual transmission better for MPG”
2. "Quantify the MPG difference between automatic and manual transmissions"

Using linear regression, we determine that the MPG for Manual transmissions is higher than that of Automatic transmission. Considering influenctial variables Manual transmission is better for MPG by a factor of 1.8 compared to Automatic transmission.

##Data Processing
We first load and review the dataset and perform necessary variables to factors
```{r, echo=FALSE}
data(mtcars)
head(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am, labels = c("Auto", "Manual"))
```

##Exploratory Analysis
In this analysis, we check effects of transmission type on MPG. We plot a boxplot of variables mpg against 
am(Figure 1). We see from the plot that, Manual transmission vehicles have a higher MPG.

##Hypothesis Testing
We can confirm the above exploratory analysis by a T-test.
```{r, echo=FALSE}
with(mtcars, t.test(mpg[am=="Auto"], mpg[am=="Manual"]))
```
T-test with p-value=0.001374 confirms that the means are significantly different.

##Regression Analysis
We first build a simple regression model.
```{r, echo=FALSE}
fit1 <- lm(mpg~am, data = mtcars)
summary(fit1)
```
Automatic transmission has lower mean than Manual transmission which agrees with the exploratory boxplot and t-test. However, the Multiple R-squared value is `r round(summary(fit1)$r.squared,4)`. This explains that the model only accounts for `r round(summary(fit1)$r.squared,4)*100`% of the variance.

We will now build a regression model including variables that also have an influence on mpg. Checking for correlations, we notice that disp and cyl are closely correlate and thus choose to ignore disp.
```{r, echo=FALSE}
data("mtcars")
sort(cor(mtcars)["mpg",])
```

```{r, echo=FALSE, results='hide'}
data(mtcars)
head(mtcars)
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am, labels = c("Auto", "Manual"))
```

We will model our regression using variable wt, am, cyl and hp.
```{r, echo=FALSE}
model <- lm(mpg~am+wt+hp+cyl, mtcars)
summary(model)
```
After including the additional variables, we can see that the new model accounts for `r round(summary(model)$r.squared,4)*100`% of the variance.

We can confirm that the models are significantly different by an ANOVA check.
```{r, echo=FALSE}
anova(fit1, model)
```
With a p-value of 1.688e-08, confirm that the new model is significantly different from our simple model.

##Residual plot and Diagnostics
In this section, we analyze the residual plots of our final regression model. Referring to Figure 2, 
we conclude that the residuals are normally distributed.

##Conclusion
With the above analysis, we conclude that, on an average, Manual Transmission cars have higher MPGs than Automatic transmission cars by a factor of 1.089. 


##Appendix
```{r, fig.cap="Figure 1",echo=FALSE}
boxplot(mpg~am, data=mtcars, xlab="Transmission", ylab="MPG", main="Transmission vs. MPG")
```


```{r, fig.cap='Figure 2', echo=FALSE}
par(mfrow=c(2,2))
plot(model)
```